---
title       : And the winner is . . .
subtitle    : Exploration of the oscar nominees
author      : Julien Beaulieu
job         : M.Eng, Polytechnique Montreal
framework   : revealjs   # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
github      :
    user: julienfbeaulieu
    repo: oscarSlidify
---

## Introduction

```{r echo=FALSE, results='hide',message=FALSE}
library(dplyr) #for filter function
library(ggplot2) # for outputting graphs

movie <- read.csv("movie.csv")
```


- Every year, [Academy Awards]("https://en.wikipedia.org/wiki/Academy_Awards") are given to the best movies featured in the previous year.
- The `Oscar shiny app` explores data from the movies that were nominated in the Oscar ceremonies.  
- The app can assessed on the [shinyapps.io]("https://jbeaulieu.shinyapps.io/oscar")
- Alternatively, the app can be run locally by:
  1. downloading the `server.R` and `ui.R` files from the [github]("https://github.com/julienfbeaulieu/oscar-project""), 
  2. setting the working directory to where the files are located,
  3. running the following R code:

```{r eval=FALSE}
library(shiny)
runApp()
```


---

## Methodology
1. The [original dataset]("https://www.aggdata.com/free_data_awards_locations/academy_awards") lists information about each nomination for each year. 
    - Since an award can either go to a film or to an artist, I processed the database to distinguish the two. The resulting `oscar.csv` database lists 10 variables about 10098 nominations.
    
2. From this, a second dataset was created and takes the point of view of each film. Number of academy nominations and awards were summed, and extra information was pulled out using the [omdb api]("http://www.omdbapi.com/"). The resulting `movie.csv` database lists 27 variables about 4361 nominated movies.
    - For instance, the following call retrieves information about the [titanic]("https://en.wikipedia.org/wiki/Titanic_(1997_film)") movie:
 
```{r eval=FALSE}
  t <- "Titanic"
  y <- "1997"
  url <- paste0("http://www.omdbapi.com/?t=",t,"&y=",y,"&plot=short&r=json")
  
  temp <- as.data.frame(fromJSON(url))
``` 

---


## Side panel

### Filter movies
In the side panel, you can filter the dataset, either by year, by number of oscar nominations or awards.
    - For instance, the data can be filtered for the period 1980-2000, with a minimum number of 5 nominations and 0 awards.
    - The movie database is the filtered using the following r code:
    
```{r}
    minYear <- 1980; maxYear <- 2000; minNomination <- 5; minAward <- 0
    movieSubset <- filter(movie, Year >= minYear, Year <= maxYear, 
                          Academy.Nominations >= minNomination, Academy.Awards >= minAward)
```
    
### Select variables
Additionnaly, you can select variables for the x- and y-axis of the graph and the for the linear model.
    - For instance, the [metascore]("https://en.wikipedia.org/wiki/Metacritic#Metascores") and the [imdb user rating]("https://en.wikipedia.org/wiki/Internet_Movie_Database") could be selected as the x and y variables.

```{r}
    plotData <- data.frame(y = movieSubset[["imdbRating"]], x = movieSubset[["Metascore"]])
```

---

## Output graph
For the preceeding filtered subset of movies and the variables selected, the Oscar shiny app would output the following graph:

```{r echo=FALSE, error=FALSE, warning=FALSE}
    p <- ggplot(plotData, aes(x, y)) + 
      geom_point(colour = "dodgerblue4") +
      xlab("Metascore") +
      ylab("imdbRating") +
      stat_smooth(method="lm", se=FALSE)
    
    print(p)
```

---

## Output linear model

- For the same filter and variable setting, the Oscar shiny app would ouput this linear model:

```{r echo=FALSE}
    fit1 <- lm(as.formula(paste("imdbRating"," ~ ","Metascore")), data = movieSubset)
    summary(fit1)
```

- Namely, for the highly nominated movies (>= 5 nominations) of the period 1980-2000, the Metascore is highly significant in predicting the imdbRating. Though, the correlation between the two variables is really low.




